---
title: "What do FEs do the spectrum?"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
pacman::p_load('tidyverse', 'fixest', 'lubridate')
theme_set(theme_bw())
set.seed(123)
if(Sys.info()['user'] == "tombearpark") {
  dir <- "/Users/tombearpark/Documents/GitHub/fourier_fe/code"
}else{
  stop("You need to set your paths up")
}
source(file.path(dir, "0_funcs.R"))
```

# Section 1: How do FEs affect the spectrum?

First, lets simulate a simple time series...

```{r}
beta <- 0.9

dates   <- 
  expand_grid(unit = 1, year = 1990:2010, month = 1:12) %>% 
  mutate(date = make_date(year = year, month = month, day = 1))

gen_ts <- function(df, newvar, beta = 0.95){
  df[newvar]   <- NA
  df[1,newvar] <- 0
  N <- length(df[[newvar]])
  for(ii in 2:N) df[ii, newvar] <- beta * df[ii - 1, newvar] + rnorm(1) 
  df %>% 
    group_by(year) %>% 
      mutate(!!newvar := .data[[newvar]] + rnorm(n = 1, sd = 2)) %>% 
    ungroup()
}
df <- gen_ts(dates, "y1")
m  <- feols(df, y1 ~ 1 | year)
df$residuals1 <- m$residuals
```

Plot the series...

```{r}
df %>% 
  pivot_longer(c(y1, residuals1)) %>% 
  ggplot() + 
  geom_line(aes(x = date, y = value, color = name), alpha = 0.9)

```

Get the spectrum of the time series. Calculate a FE

```{r}
compare_spectrums <- function(df, ts1, ts2){
  unit <- df$unit
  ss <- spectrum(df[[ts1]], log = "no", plot = FALSE)
  # Get the spectrum of the residuals
  s1 <- spectrum(df[[ts2]], log = "no", plot = FALSE)
  # Organise and plot outputs...
  tibble(freq = ss$freq, y = ss$spec, filtered = s1$spec) %>% 
    pivot_longer(cols = -c(freq))
}

out <- compare_spectrums(df, "y1", "residuals1")

ggplot(out) + geom_line(aes(x = freq, y = value, color = name))
```

Clearly, for an individual time series, filtering the data in this way eliminates the low frequency information.

Next, lets expand this to a panel data setting.

```{r}

gen_panel <- function(no_units){
  map_dfr(1:no_units, 
          function(x) 
            expand_grid(year = 1990:2010, month = 1:12) %>% 
              mutate(date = make_date(year = year, month = month, day = 1), 
                     unit = !!as.factor(x)) %>% 
            gen_ts("y")
  )
}

df2 <- gen_panel(5)

ggplot(df2) + geom_line(aes(x = date, y = y, color = unit))

```

```{r}
m2 <- feols(y ~ 1 | year + unit, data = df2)
df2$residuals <- m2$residuals

m3 <- feols(y ~ 1 | year^unit, data = df2)
df2$residuals_full <- m3$residuals


# Get the spectra of each unit 
compare_spectrums_by_unit <- function(df, ts1, ts2){
  map_dfr(unique(df$unit), 
        function(x){
          df %>% filter(unit == !!x) %>% 
            compare_spectrums(ts1, ts2) %>% 
            mutate(unit = !!x)
        })
}

plot_panel <- function(df){
    ggplot(df) + 
      geom_line(aes(x = freq, y = value, color = name)) + 
      facet_wrap(~unit, scales = "free")
}
compare_spectrums_by_unit(df2, "y", "residuals") %>% 
  plot_panel()


```

```{r}
compare_spectrums_by_unit(df2, "y", "residuals_full") %>% 
  plot_panel()

```

# Section 2: Getting the same coefficients using FE and Fourier

```{r}
dates   <- 
  expand_grid(unit = 1, year = 1980:2010, month = 1:12) %>% 
  mutate(date = make_date(year = year, month = month, day = 1))

df <- gen_ts(dates, "y", beta = 0.8)
df$ly <- dplyr::lag(df$y)

# Estimate using FE
m  <- feols(y ~ ly | year, data = df)
coef(m)
```

```{r}
Q <- 2 * dim(df)[1] / 12
df <- add_fitted_vals(df, Q, "y")
ggplot(df) + 
  geom_line(aes(x = date, y = y), color = "red", alpha = 0.5) +
  geom_line(aes(x = date, y = residual)) + 
  geom_line(aes(x = date, y = yhat)) + 
  ggtitle("Red line is sum of the two black lines")
```

```{r}

lm(data = df, residual ~ dplyr::lag(residual))
```
