---
title: "Orthonormal Matrices"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```
A square matrix $Q$ is orthonormal if $Q'Q=I$. Let $\mathbf{v}_j$ be the $j$th column of $Q$, then, the definition implies that $\mathbf{v}_i'\mathbf{v}_i=1, i=1,\ldots,n$ and $\mathbf{v}_i'\mathbf{v}_j=0,i\neq j$, that is that the columns of $Q$ are all orthogonal to each other, i.e. the inner product of any two columns is 0.

Say we have data $(X,y)$. Rotating the data by $Q$ does not affect our least squares estimator $\widehat{\beta}$. Note that
\[\widehat{\beta} = (X'X)^{-1}X'y=(X'Q'QX)^{-1}X'Q'Qy=(\widetilde X'\widetilde X)^{-1}\widetilde{X}'\widetilde{y},\]
where $\widetilde X=QX$ and $\widetilde y = Qy$. More is true. Note that 
\[Q'Q = I\implies Q'QQ^{-1}=Q^{-1}\implies Q'=Q^{-1}\]
thus
\[\widehat{\beta} = (X'X)^{-1}X'y=(X'QQ^{-1}X)^{-1}X'QQ^{-1}y=(X'QQ''X)^{-1}X'QQ'y=(\widehat X'\widehat X)^{-1}\widehat{X}'\widehat{y},\]
Let's see a practical example now.

```{r}
data <- cars
Q <- randortho(length(cars$speed))
data$speedQ <- Q %*% data$speed
data$distQ <- Q %*% data$dist
data$speedQtr <- t(Q) %*% data$speed
data$distQtr <- t(Q) %*% data$dist
```
Regression with raw data.
```{r}
lm(dist ~ speed - 1 , data = cars)
```
Regression with data pre-multiplied by $Q$.
```{r}
lm(distQ ~ speedQ - 1 , data = data)
```
Regression with data pre-multiplied by $Q'$.
```{r}
lm(distQtr ~ speedQtr - 1 , data = data)
```
